workshop

### 2023-02-15 scripts/workshop/dynamics.Rmd
make script for numerical solutions to ODE's

---
title: "alternate fig 1"
author: "Daniel Suh"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

@@

```{r, message = F}
library(tidyverse)
library(magrittr)
#install_github("thomasp85/patchwork")
library(patchwork)
library(ggnewscale)

@@

  data$span1 <- 1/data$mort1 #lifespan for species 1
  data$span2 <- 1/data$mort2 #lifespan for species 2
  data$half <- log(1/2)/-data$degr #viral half-life
  data$epid <- data$eigen > 1 #invasion potential (i.e. R0 > 1)
  data$prop1 <- data$trans1/data$trans2 #proportion of transmission of species 2 over species 1
  data$prop3 <- data$trans3/data$trans2 #proportion of environmental transmission over species 1
  data$prop_mort <- data$mort1/data$mort2 #relative mortality for species 1 and 2
  data$cap1 <- data$birth1/data$mort1 #carrying capacity for species 1
  data$cap2 <- data$birth2/data$mort2 #carrying capacity for species 2
  data$cap_prop <- data$cap1/data$cap2 #relative carrying capacities for species 1 and 2
  data$tot <- data$cap1 + data$cap2 #total abundance for species 1 and 2
  data$cc <- data$trans1*(data$cap1/data$tot) + data$trans2*(data$cap2/data$tot) #cc defined as transmission of a species multiplied by its relative abundance
  output <- data %>% dplyr::select(trans1, trans2, prop1, prop3, trans3, eigen, prop_mort, mort1, mort2, degr, half, cap1, cap2, cap_prop, tot, cc) %>% distinct()
  return(output)
}
```

```{r}
reference <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                     mort1 = 1/45,mort2 = 1/45,degr = 1/1.947799)

composition <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                       mort1 = 1/60,mort2 = 1/30,degr = 1/1.947799)

size <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                mort1 = c(1/52.5),mort2 = c(1/52.5),degr = 1/1.947799)

halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                    mort1 = c(1/40),mort2 = 1/40,degr = 1/3.895598)

combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0,trans_3_max = 0.001,
                    mort1 = c(1/70),mort2 = c(1/35),degr = 1/3.895598)
```

```{r}
axis_text_size = 8
plot_label_size = 20

p1 <- reference %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(na.value = NA, values = c(NA, "red")) + 
  labs(x = "Environmental", y = "Contact") +
  xlim(1,8) +
  ylim(1,14) + 
  labs(title = "Reference")

p2 <- composition %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(na.value = NA, values = c(NA, "red")) + 
  new_scale_color()+
  geom_contour(data = reference, 
               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
  labs(x = "", y = "") +
  xlim(1,8) +
  ylim(1,14) + 
  labs(title = "A") + 
  theme(axis.title = element_text(size=axis_text_size),
        plot.title = element_text(size=plot_label_size))
#p2

p3 <- size %>% filter(tot != 150) %>% 
  ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  theme_classic()+
  scale_colour_manual(na.value = NA, values = c(NA, "orange")) + 
  new_scale_color()+
  geom_contour(data = reference, 
               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
  labs(x = "", y = "") +
  xlim(1,8) +
  ylim(1,14) + 
  labs(title = "B") + 
  theme(plot.title = element_text(size=plot_label_size))
#p3

p4 <- halflife %>% ggplot(.,aes(x=prop3,y=prop1))+
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE)+
  scale_colour_manual(na.value = NA, values = c(NA, "forestgreen")) + 
  new_scale_color()+
  geom_contour(data = reference, 
               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F)+
  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
  labs(x="Relative Environmental", y = "                                                               Relative Contact Transmission Rate", title = "C") +
  xlim(1,8) +
  ylim(1,14) + 
  theme_classic() +
  theme(axis.title = element_text(size=axis_text_size),
        plot.title = element_text(size=plot_label_size),
        axis.title.x=element_text(hjust=1.05))
#p4

p5 <- combined %>%
  ggplot(.,aes(x=prop3,y=prop1)) +
  geom_contour(aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = FALSE) +
  scale_colour_manual(na.value = NA, values = c(NA, "purple")) + 
  new_scale_color()+
  geom_contour(data = reference, 
               aes(x=prop3,y=prop1,z=eigen,colour = factor(..level.. == 1,levels = c(F,T)),group=mort1), show.legend = F) +
  scale_colour_manual(na.value = NA, values = c(NA, "gray")) +
  labs(x = "Transmission Rate", y = "", title = "D") +
  xlim(1,8) +
  ylim(1,14) + 
  theme_classic() +
  theme(axis.title = element_text(size=axis_text_size),
        plot.title = element_text(size=plot_label_size),
        axis.title.x=element_text(hjust=-0.625))
#p5
```

```{r}
figure_1 <- (p2 | p3)/(p4 | p5)
figure_1 + theme(title = "Original fig w/ lines for R0==1")
```

```{r}
#heatmaps and contour plots of difference between control and treatment

ref_sel <- reference %>% select(prop3,prop1,eigen) %>% rename(.,ref_eigen = eigen)
comp_sel <- composition %>% select(prop3,prop1,eigen) %>% rename(.,comp_eigen = eigen)
size_sel <- size %>% select(prop3,prop1,eigen) %>% rename(.,size_eigen = eigen)
halflife_sel <- halflife %>% select(prop3,prop1,eigen) %>% rename(.,half_eigen = eigen)
combined_sel <- combined %>% select(prop3,prop1,eigen) %>% rename(.,all_eigen = eigen)

ref_comp <- inner_join(ref_sel,comp_sel, by = c("prop3","prop1"))
ref_comp %<>% inner_join(.,size_sel, by =c("prop3","prop1"))
ref_comp %<>% inner_join(.,halflife_sel, by =c("prop3","prop1"))
ref_comp %<>% inner_join(.,combined_sel, by =c("prop3","prop1"))

ref_comp %<>% mutate(.,comp_diff = comp_eigen-ref_eigen,
                     size_diff = size_eigen-ref_eigen,
                     half_diff = half_eigen-ref_eigen,
                     all_diff = all_eigen-ref_eigen)

#heatmaps with contours

comp_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=comp_diff)) + 
  geom_tile() +
  scale_fill_gradient2() + 
  geom_contour(aes(z=comp_diff))

size_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=size_diff)) + 
  geom_tile() +
  scale_fill_gradient2() + 
  geom_contour(aes(z=size_diff))

half_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=half_diff)) + 
  geom_tile() +
  scale_fill_gradient2() + 
  geom_contour(aes(z=half_diff))

combined_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=all_diff)) + 
  geom_tile() +
  scale_fill_gradient2() + 
  geom_contour(aes(z=all_diff))
```

```{r}
(comp_fig | size_fig)/(half_fig | combined_fig) + theme(title = "heatmap of difference between treatment and reference")
```

```{r}
#filled contour with standardized breaks

breaks <- c(-0.2,-0.1,0,0.1,0.2,0.3,0.4,0.5,2)

comp_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=comp_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +
  theme(legend.position = "none") + 
  labs(x="environmental transmission", y = "contact transmission", title = "composition")

size_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=size_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +
  theme(legend.position = "none") + 
  labs(x="environmental transmission", y = "contact transmission", title = "abundance")

half_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=half_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +
  theme(legend.position = "none") + 
  labs(x="environmental transmission", y = "contact transmission", title = "halflife")

combined_fig <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=all_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +   
  labs(x="environmental transmission", y = "contact transmission", title = "combined")
```

```{r}
(comp_fig | size_fig)/(half_fig | combined_fig) + theme("filled contours of diff between treatment and reference")
```

```{r}
#comparison between combined treatment and the additive effect

ref_comp %<>% rowwise %>% mutate(diff_sum = comp_diff + size_diff + half_diff)
ref_comp %<>% rowwise %>% mutate(effect = all_diff-diff_sum)

# ref_comp %>% ggplot(.,aes(x=prop3,y=prop1,fill=effect)) + 
#   geom_tile() +
#   scale_fill_gradient2() + geom_contour(aes(z=effect))
# 
# ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
#   geom_contour_filled(aes(z=effect))

breaks <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9)

comb <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=all_diff),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +   
  theme(legend.position = "none") + 
  labs(x="", y = "", title = "combined")

sum <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=diff_sum),breaks = breaks) +
  scale_fill_viridis_d(drop=F) + 
  labs(x = "", y = "contact transmission", title = "sum of its parts")

diff <- ref_comp %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=effect),breaks = breaks) +
  scale_fill_viridis_d(drop=F) + 
  theme(legend.position = "none") +
  labs(x = "environmental", y = "", title = "difference (combined - sum of its parts)")

comb / sum / diff 
```

```{r}
#comparison between combined treament and additive effect by percent change

ref_percent <- ref_comp %>% mutate(., comp_per = comp_diff/ref_eigen,
                                   size_per = size_diff/ref_eigen,
                                   half_per = half_diff/ref_eigen,
                                   all_per = all_eigen/ref_eigen,
                                   sum_per = diff_sum/ref_eigen)

breaks <- c(0:20)*0.1

comb_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=all_per),breaks = breaks) +
  scale_fill_viridis_d(drop=F) +   
  theme(legend.position = "none") + 
  labs(x="", y = "", title = "combined")

sum_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=sum_per),breaks = breaks) +
  scale_fill_viridis_d(drop=F) + 
  labs(x = "", y = "contact transmission", title = "sum of its parts")

diff_per <- ref_percent %>% ggplot(.,aes(x=prop3,y=prop1)) + 
  geom_contour_filled(aes(z=all_per-sum_per),breaks = breaks) +
  scale_fill_viridis_d(drop=F) + 
  theme(legend.position = "none") +
  labs(x = "environmental", y = "", title = "difference (combined - sum of its parts)")

comb_per / sum_per / diff_per 
```

>>>>

---
title: "dynamics"
author: "Daniel Suh"
date: "2023-02-15"
output: html_document
---

@@

```{r}
library(tidyverse)
library(magrittr)
#install_github("thomasp85/patchwork")
library(patchwork)
library(ggnewscale)
library(here)
```

@@

  data$span1 <- 1/data$mort1 #lifespan for species 1
  data$span2 <- 1/data$mort2 #lifespan for species 2
  data$half <- log(1/2)/-data$degr #viral half-life
  data$epid <- data$eigen > 1 #invasion potential (i.e. R0 > 1)
  data$prop1 <- data$trans1/data$trans2 #proportion of transmission of species 2 over species 1
  data$prop3 <- data$trans3/data$trans2 #proportion of environmental transmission over species 1
  data$prop_mort <- data$mort1/data$mort2 #relative mortality for species 1 and 2
  data$cap1 <- data$birth1/data$mort1 #carrying capacity for species 1
  data$cap2 <- data$birth2/data$mort2 #carrying capacity for species 2
  data$cap_prop <- data$cap1/data$cap2 #relative carrying capacities for species 1 and 2
  data$tot <- data$cap1 + data$cap2 #total abundance for species 1 and 2
  data$cc <- data$trans1*(data$cap1/data$tot) + data$trans2*(data$cap2/data$tot) #cc defined as transmission of a species multiplied by its relative abundance
  output <- data %>% dplyr::select(trans1, trans2, prop1, prop3, trans3, eigen, prop_mort, mort1, mort2, degr, half, cap1, cap2, cap_prop, tot, cc) %>% distinct()
  return(output)
}

reference <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/45,mort2 = 1/45,degr = 1/1.947799)

composition <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = 1/60,mort2 = 1/30,degr = 1/1.947799)

size <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/52.5),mort2 = c(1/52.5),degr = 1/1.947799)

halflife <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/40),mort2 = 1/40,degr = 1/3.895598)

combined <- contour(trans_1_min = 0.0001,trans_1_max = 0.001,trans_3_min = 0.0001,trans_3_max = 0.001,mort1 = c(1/70),mort2 = c(1/35),degr = 1/3.895598)

### 2023-02-15 scripts/workshop/workshop.Rmd
make script for numerical solutions to ODE's

effect sizes
```{r}
comm_data <- readRDS(here("processed_data","comm_data.rds"))
comm_data %>% mutate(mean = mean(size), sd = sd(size), var = mean(size)-size, size = size)
comm_effect <- comm_data %>% filter(MeanWaterTempPredC>0) %>% mutate(size_sds = (mean(size)-size)/sd(size),
                                    cc_sds = (mean(cc)-cc)/sd(cc),
                                    temp_sds = ((mean(MeanWaterTempPredC)-MeanWaterTempPredC)/sd(MeanWaterTempPredC))*-1) #multiply by -1 to make it opposite because

>>>>

Standard deviations from the mean
This chunk attempts to characterize each wetland in terms of how exceptional they are compared to other wetlands in terms of abundance, cc, and temp
The idea is that if certain wetlands tend to be highly abundant with high community competence and lower temperatures, then we might expect them to be a perfect environment for ranavirus to thrive. 
```{r}
comm_data <- readRDS(here("processed_data","comm_data.rds"))
comm_data %>% mutate(mean = mean(size), sd = sd(size), var = mean(size)-size, size = size)
comm_effect <- comm_data %>% filter(MeanWaterTempPredC>0) %>% mutate(size_sds = (mean(size)-size)/sd(size),
                                    cc_sds = (mean(cc)-cc)/sd(cc),
                                    temp_sds = ((mean(MeanWaterTempPredC)-MeanWaterTempPredC)/sd(MeanWaterTempPredC))*-1) #multiply by -1 to make it opposite because
